---
title: "EAS648 Lab05"
author: "Haolin Li (haolinli@umich.edu)"
date: "2023-11-16"
output: 
  html_document:
    output_file: "index.html"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
Sys.setenv(LANG = "en")
setwd("D:/UMich/Fall23/EAS648/Repo/Lab/05")
```

# EAS 648 Lab 05: Applied spatial analysis

> **Links: **
>
> **Course Repository**: https://github.com/HumblePasty/EAS648
>
> **Lab05 Webpage**: https://humblepasty.github.io/EAS648/Lab/05/
>
> **Lab05 Repository**: https://github.com/HumblePasty/EAS648/tree/master/Lab/05


> Assignment
>
> 1. Analyze the Salt Lake City dataset. Fit a model that explains suitability for urban development (provide evidence of goodness of model fit). What characteristics contribute to urban development in Salt Lake? What variable might you add to better explain urban development. (1-2 paragraphs)
> 2. Analyze the database of amenity landscape characteristics using a clustering technique. Choose variables based on a thematic inquiry of your choice. Visualize ad provide qualitative descriptions of the groups that you derive from the clustering techniques. Provide evidence of you characterizations. Provide a map of your cluster analysis. Give an explanation of the spatial distribution of these groups. (1-2 paragraphs)



> Haolin's Note:
>
> For this assignment, I'll complete the tasks in 2 parts:
>
> - Part I: Analyze the Salt Lake City dataset and Model Fitting
> - Part II: Cluster Analysis of Landscape Characteristics

## Part I: Analyze the Salt Lake City dataset and Model Fitting

### Preparing the Dataset

```{r}
library(terra)

#first import the land cover/use data
NLCD_2001 <- rast("Lab5prj/NLCD_2001_SL.tif")
NLCD_2004 <- rast("Lab5prj/NLCD_2004_SL.tif")
NLCD_2006 <- rast("Lab5prj/NLCD_2006_SL.tif")
NLCD_2008 <- rast("Lab5prj/NLCD_2008_SL.tif")
NLCD_2011 <- rast("Lab5prj/NLCD_2011_SL.tif")
NLCD_2013 <- rast("Lab5prj/NLCD_2013_SL.tif")
NLCD_2016 <- rast("Lab5prj/NLCD_2016_SL.tif")
# Distance to parks and protected areas in km (Euclidean) for the study area
Park_dist <- rast("Lab5prj/Parks_dist_SL.tif")
# Road density for a 1 km neighborhood
Rd_dns1km <- rast("Lab5prj/Rd_dns1km_SL.tif")
# Distance to water bodies in km (Euclidean)
WaterDist <- rast("Lab5prj/WaterDist_SL.tif")
# Elevation
DEM <- rast("Lab5prj/DEM_SL.tif")

allrasters <- c(NLCD_2001, NLCD_2004, NLCD_2006, NLCD_2008, NLCD_2011, NLCD_2013, NLCD_2016, Park_dist, Rd_dns1km, WaterDist,DEM)

library(tidyverse)
allrastersSL <- as.data.frame(allrasters, xy=TRUE)
## Here we are filtering out the no data values (stored as 128)
allrastersSL <- allrastersSL %>%
  filter (NLCD_2001_SL != 128)
head(allrastersSL)
```

> Pre-plotting the Data

```{r}
plot(NLCD_2001)
```


### Sampling the Data

```{r}
library(leaflet)
sampleSLrnd <- spatSample(allrasters, size=100, "random", cells=TRUE, xy=TRUE)
head(sampleSLrnd)
plot(sampleSLrnd$x, sampleSLrnd$y, main = "Sample Locations for Autocorrelation Analysis")
```

### Spatial Autocorrelation Analysis (Moran's I)

```{r}
# flatten the spatial data to a dataframe that has lat and long
flat_data <- as.data.frame(sampleSLrnd)
flat_data = na.omit(flat_data)
# calculate distances between all the points,    
dist_matrix <- as.matrix(dist(cbind(flat_data$x, flat_data$y)))
# and generate a matrix of inverse distance weights.
dist_matix.inv <- 1/dist_matrix
diag(dist_matix.inv) <- 0

library(ape)
# Moran.I(sampleSLrnd$Rd_dns1km_SL, dist_matix.inv)
Moran.I(flat_data$Rd_dns1km_SL, dist_matix.inv)
```

### Plotting and Statistical Analysis of the Changed Area

```{r}
allrastersSL <- allrastersSL %>%
    mutate(urbanChg = (NLCD_2001_SL != 21 & NLCD_2001_SL != 22 & NLCD_2001_SL != 23 & NLCD_2001_SL != 24) &  (NLCD_2016_SL == 21 | NLCD_2016_SL == 22  | NLCD_2016_SL == 23 | NLCD_2016_SL == 24))

ggplot(allrastersSL, aes(y=y, x=x, color=urbanChg)) +
   geom_point(size=2, shape=15) +
   ggtitle("Changed Urban Areas Distribution between 2001 and 2016") +
   theme()
```

```{r}
## calculate total new urban impervious for 2016
newUrban <- (sum(as.numeric(allrastersSL$NLCD_2016_SL == 21 | allrastersSL$NLCD_2016_SL == 22 |allrastersSL$NLCD_2016_SL == 23 | allrastersSL$NLCD_2016_SL == 24))) - (sum(as.numeric(allrastersSL$NLCD_2001_SL == 21| allrastersSL$NLCD_2001_SL == 22| allrastersSL$NLCD_2001_SL == 23| allrastersSL$NLCD_2001_SL == 24)))
## calculate total urban impervious for 2001
urban2001 <- (sum(as.numeric(allrastersSL$NLCD_2001_SL == 21| allrastersSL$NLCD_2001_SL == 22| allrastersSL$NLCD_2001_SL == 23| allrastersSL$NLCD_2001_SL == 24)))
## percentage increase in urban impervious
newUrban/urban2001* 100
```

```{r}
library(plyr)
library(reshape2)
library(plotly)

SL <- allrastersSL %>%
  # the non urban areas in 2001
  filter(NLCD_2001_SL != 21 & NLCD_2001_SL != 22 & NLCD_2001_SL != 23 & NLCD_2001_SL != 24) 
SL <- SL[10:14]
# similar to tidy(), melt is a function to organize the data
SLmelt<-melt(SL)
p <- ggplot(SLmelt, aes(x=urbanChg, y=value,fill=variable))+
    geom_boxplot()+
    facet_grid(.~variable)+
    labs(x="X (binned)")+
    theme(axis.text.x=element_text(angle=-90, vjust=0.4,hjust=1))
p
```


### Fitting the Model


```{r}
###Grab all the developed cells (presence)
newUrban <- SL %>%
  filter(urbanChg == TRUE)

###Grab all the nondeveloped and not previously urban cells (absence)
nonUrban <- SL %>%
  filter(urbanChg == FALSE)

###Get a random sample of the absence data  
### that is twice as large as the presence data
index <- sample(1:nrow(nonUrban), (round(nrow(newUrban)* 2)))
SLsampleUrban <- nonUrban[index, ]

### combine the original presence and absence data
SLsample <- rbind(SLsampleUrban, newUrban)

###Consider making a training and testing dataset
###This can reduce the computational burden 
### It also is a robust goodness of fit method
SLsample <- SLsample %>% dplyr::mutate(id = row_number())
#Create training set
train <- SLsample %>% sample_frac(.70)
#Create test set
test  <- anti_join(SLsample, train, by = 'id')
```

> Fit a Model

```{r}
fit <- glm(urbanChg ~ Parks_dist_SL + Rd_dns1km_SL + WaterDist_SL + DEM_SL,data=train,family=binomial())
summary(fit)
```

### Test the Model

> Test the data

```{r}
## Loading required package: ggplots
library(ROCR)
# plot a ROC curve for a single prediction run
# and color the curve according to cutoff.
pred <- prediction(predict(fit, newdata = test), test$urbanChg)
perf <- performance(pred,"tpr","fpr")
plot(perf,colorize=TRUE)
```

```{r}
## A simpler way to understand these result is to calculate the
## area under the curve(AUC). The closer this number is to 1, the
## better your model fit
auc_ROCR <- performance(pred, measure = "auc")
auc_ROCR <- auc_ROCR@y.values[[1]]
auc_ROCR
```

### Predict the Result Using the Model

```{r}
predicted = predict(allrasters, fit)
plot(predicted)
```



> **Analysing the Fitting Goodness**
>
> The goodness of the fitting result can be assessed by two aspects:
>
> 1. As shown by the ``summary(fit)`` result, all the explanatory variables are significant for the explaining the changed areas.
> 2. The ROC line is steep at the beginning and AUC is close to 1. Thus the model is good in fitting the result.
>
> **What characteristics contribute to urban development in Salt Lake?**
>
> According to the coefficients, distance to road has the largest value, which implies that distance to roads (which means transportation convenience) is the most significant (positive) factor that may lead to development. Distance to parks and distance to water have much smaller coefficients. The coefficent is negative, which can be intepret as "higher lands may require more effort to develop".
>
> **What variable might you add to better explain urban development**
>
> - **Normalized Difference Vegetation Index (NDVI)**: This index, derived from satellite imagery, measures vegetation health and coverage. It can be used to monitor green spaces within urban areas.
> - **Air Quality Indices**: Using satellite data, it’s possible to estimate the concentration of different pollutants, providing insights into environmental aspects of urban areas.
> - **Nighttime Lights**: Satellite imagery of nighttime lights is often used as a proxy for urbanization and economic activity.
> - **Land Surface Temperature**: Using thermal imagery, land surface temperature can be a good indicator of urban heat islands and overall urbanization levels.
> - ...



## Part II: Cluster Analysis of Landscape Characteristics

### Loading the Data

```{r}
library(fastcluster)
library(graphics)
library(ggplot2)
library(ggdendro)

library(sf)
amenData<- st_read("Lab5prj/AmenDataAll.shp")
# head(amenData)
```

```{r}
geomData = st_sfc(amenData$geometry)
```

### Pre-Process

```{r}
amenData$ZooEmpDist_log <- log(amenData$ZooDistEmp + 0.00000001)
amenData$HotelEmpDist_log <- log(amenData$HotelDistE + 0.00000001)
amenData$HistMonDist_log <- log(amenData$HistMon_Di + 0.00000001)
amenData$HighEdEmpDist_log <- log(amenData$HigherEdDi + 0.00000001)
amenData$GolfEmpDist_log <- log(amenData$GolfDistEm + 0.00000001)

amenData$SocialNorm <- amenData$Nat_Flickr/(amenData$serPop10 + 1)
amenData$HousingChg <- amenData$Urb2011 - amenData$Urb2001

amenDataDF<-amenData[,c("SocialNorm", "HousingChg", "Frst2011", "WaterPct", "distcoast", "DEM_max","DEM_range", "HikeLength","ZooEmpDist_log", "HotelEmpDist_log","HistMonDist_log", "HighEdEmpDist_log", "GolfEmpDist_log")]
###if you want to add variable look at the names using: names(amenData)

## make sure there are no missing data
amenDataDF <- na.omit(amenDataDF)
## we need to make it into a normal dataframe to 
## do the analysis
amenDataDF <- as.data.frame(amenDataDF)[1:12]
## calculate z-scores for the dataset
db_scaled <- scale(amenDataDF)
```

### K-means clustering

> Determine the number of clusters

```{r}
# Determine number of clusters
wss <- (nrow(db_scaled)-1)*sum(apply(db_scaled,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(db_scaled,
   centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
  ylab="Within groups sum of squares")
```
> Cluster

```{r}
# K-Means Cluster Analysis
fit_new <- kmeans(db_scaled, 5) # 5 cluster solution

# get cluster means
aggregate(db_scaled,by=list(fit_new$cluster),FUN=mean)
```

> Plot the Data

```{r}
# append cluster assignment
amenData <- data.frame(amenData, fit_new$cluster)

st_geometry(amenData) <- geomData

ggplot() + 
  geom_sf(data = amenData, mapping = aes(fill = as.factor(fit_new.cluster)), color = NA) + 
  theme(legend.position = "none") +
  ggtitle("Clusters based on Kmeans")
```

> TBD: rechoose the variables, add compass, north arrow, scale bar

> Descriptive questions
